---
title: 'results'
author: 'garezana'
output: pdf_document
---

```{r, include=FALSE}

library(ggpubr)
library(ggplot2)
library(dplyr)
library(latex2exp)
library(gridExtra)
library(lubridate)
library(ggnewscale)
```

##### Note:Data AND Rmd file should be in same folder

```{r WIV dataframe, include=FALSE}
#Importing data set
erythrocyteMatureSet = read.csv('gundMatureErys.csv') # Mature erythrocyte summary data
erythrocyteImmatureSet = read.csv('gundImmatErys.csv') # Immature erythrocyte summary data
seasonalGundlachiAzuroSet = read.csv('seasonalGundAzuro_dat1.csv') # Seasonal and individual data
```

```{r seasonaldataformat, include=FALSE}
# Creating 'Season' and 'infection.status' features, formatting 'Date' to have date format and selecting wanted features
seasonalGundlachiAzuroDF = seasonalGundlachiAzuroSet %>%
  dplyr::mutate (
    Date = as.Date(Date, format = '%m/%d/%y'), 
    Season = case_when (month(Date) %in% c(6, 7 ,8) ~ 'summer',
                       (month(Date) %in% c(1, 2, 3)) ~ 'winter' ),
    infection.status = as.factor(if_else(AZR..Slide. == 1 & Infecciones.PCR == 1, 1,
                                 if_else(AZR..Slide. == 0 & Infecciones.PCR == 0, 0, NA_real_))) 
    ) %>%
  dplyr::select(Date, Season, Spp, UID, Sex, SVL, Weight, infection.status)
```

```{r data merge, include=FALSE}
# Merging lizard and seasonal data with erythrocyte to mature erythrocyte summary data
matureErythrocyteSeasonalDF = merge(erythrocyteMatureSet, seasonalGundlachiAzuroDF, by = 'UID' ) %>% 
  select(Date, Season, UID, Spp, Sex, SVL, Weight, infection.status,
         classification, ratio, cellNum, eryMean, erySD, CV)

# Creating ordinal lizard size feature
matureErythrocyteSeasonalDF = matureErythrocyteSeasonalDF %>%
  dplyr::mutate(
    Size = ifelse(Sex == 'F' & SVL <= median(SVL[Sex == 'F'])|
                  Sex == 'M' & SVL <= median(SVL[Sex == 'M']), 'Small', 
           ifelse(Sex == 'F' & SVL >  median(SVL[Sex == 'F'])|
                  Sex == 'M' & SVL >  median(SVL[Sex == 'M']), 'Large', NA)))

```

```{r datasave, include=FALSE}
# Fix data transcription error in lizard mass
matureErythrocyteSeasonalDF$Weight[matureErythrocyteSeasonalDF$UID == '1250'] = 9.0 # Wrongly transcribed as 4
matureErythrocyteSeasonalDF$Weight[matureErythrocyteSeasonalDF$UID == '2183'] = 7.7 # Wrongly transcribed as 2.7

agundlachiWIVc = matureErythrocyteSeasonalDF %>% select(uid = UID, spp = Spp, sex = Sex, mass = Weight, svl = SVL)
write.csv(agundlachiWIVc, 'agundlachiWIV-09-2021.csv', row.names = F)
```

```{r BCI Calc, include=FALSE}
# Creating BCI feature from Weight and SVL data
matureErythrocyteSeasonalDF = matureErythrocyteSeasonalDF %>%
  dplyr::mutate(BCI = residuals(lm(log10(Weight) ~ log10(SVL))))
```

### Model1. BCI predicted by infection status, sex and season
```{r, echo=FALSE}
# Training model for BCI prediction
BCIModel = glm(BCI ~ infection.status + Sex + Season, data = matureErythrocyteSeasonalDF)
summary(BCIModel)
```

```{r, include=FALSE}

dataFrameColumnTitles = c('infection.status', 'Sex', 'Season', 'Predicted', 'lower', 'upper', 'se.predict')

# Function to create data frame with model predictions
getModelPredictionFor <- function(season) {
  dataFrameBuffer = setNames(data.frame(matrix(ncol = 7, nrow = 0)), dataFrameColumnTitles)

  # Build data frame with model predictions
  dataFrameBuffer <- rbind(dataFrameBuffer, dataPredictor('0', 'F', season))
  dataFrameBuffer <- rbind(dataFrameBuffer, dataPredictor('1', 'F', season))
  dataFrameBuffer <- rbind(dataFrameBuffer, dataPredictor('0', 'M', season))
  dataFrameBuffer <- rbind(dataFrameBuffer, dataPredictor('1', 'M', season))
  
  # Return dataframe with predictions
  return(dataFrameBuffer)
}

# Function to predict BCI using model created
dataPredictor <- function(infectionStatus, sex, season) {

  dataFrameBuffer = setNames(data.frame(matrix(ncol = 7, nrow = 0)), dataFrameColumnTitles)

  # Changing elements in data frame for predictions
  temporaryDataFrame = data.frame(infection.status = infectionStatus, Sex = sex, Season = season)
  # Predicted data for data frame
  predictedData=predict(BCIModel, newdata = temporaryDataFrame, interval = 'confidence', se.fit = T)

  lower = predictedData$fit - 1.96 * predictedData$se.fit
  upper = predictedData$fit + 1.96 * predictedData$se.fit

  # Create a list with prediction information in expected format to attach into data frame
  formatData = list(infectionStatus, sex, season, predictedData$fit, lower, upper, predictedData$se.fit)

  # Adding values to new data frame to create plots
  dataFrameBuffer[nrow(dataFrameBuffer) + 1,] = formatData
  return(dataFrameBuffer)
}
```

```{r, echo=FALSE}
# Data frames with BCI prediction
# For winter lizards
winterAPredictionsDF = getModelPredictionFor('winter')
# For summer lizards
summerAPredictionsDF = getModelPredictionFor('summer')
```

```{r, echo=FALSE}
# Binding winter and summer predicted BCI data frames
seasonAPredictionsDF = rbind(summerAPredictionsDF, winterAPredictionsDF)
knitr::kable(seasonAPredictionsDF, caption = 'Predicted body condition index (BCI) values by infection status')
```

### Figure 1 
Predicted partial relationships between infection status and body condition index (BCI) for female and male *A.gundlachi* during the A) 2015 summer and B) 2016 winter season. Grey error bars represent 95 % CI and black error bars represent one standard error.

```{r, echo=FALSE}
# Plotting winter BCI predictions
leyendText = expression(paste('Winter', italic(' Anolis gundlachi')))
winterBCIplot = ggplot(winterAPredictionsDF,
                        aes(Sex, Predicted, factor = infection.status)) +
                        geom_errorbar(position = position_dodge(.9), aes(ymin = lower, ymax  = upper),
                                      width = 0.9, color = 'gray83', size = 2) +
                        geom_errorbar(position = position_dodge(.9), aes(ymin = Predicted-se.predict, ymax = Predicted + se.predict),
                                      width = 0.5, color = 'black', size = 2) +
                        geom_point(position = position_dodge(.9), size = 3, aes(colour = infection.status)) +
                        labs(y = 'Body Condition Index') +
                        scale_color_manual(labels = c('Non-Infected', 'Infected'), values = c('#0000FF', '#f09d1d'),
                                           name = leyendText) +
                        theme_bw() +
                        theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14),
                              legend.text = element_text(size = 14), legend.title = element_text(size = 16),
                              legend.text.align = 0, axis.title.y = element_blank()) +
                        scale_x_discrete(labels = c('Female', 'Male')) +
                        theme(legend.title = element_blank()) +
                        theme(legend.position = c(0.325, 0.9))
winterBCIplot2 = winterBCIplot + annotate('text',  x = Inf, y = Inf, label = 'B', vjust = 1.5, hjust = 1.5) +
                                 scale_y_continuous(breaks = round(c(seq(-.075, 0.050, 0.025)), 3),
                                                    limits = c(-.060, 0.050))
```


```{r, echo=FALSE}
# Plotting summer BCI predictions
leyendText = expression(paste('Summer', italic(' Anolis gundlachi')))
summerBCIplot = ggplot(summerAPredictionsDF,
                          aes(Sex, Predicted, factor = infection.status)) +
                          geom_errorbar(position = position_dodge(.9), aes(ymin = lower, ymax = upper),
                                        width = 0.9, color = 'gray83', size = 2) +
                          geom_errorbar(position = position_dodge(.9), aes(ymin = Predicted-se.predict, ymax = Predicted + se.predict),
                                        width = 0.5, color = 'black', size = 2) +
                          geom_point(position = position_dodge(.9), size = 3, aes(colour = infection.status)) +
                          labs(y = 'Body Condition Index') + 
                          scale_color_manual(labels = c('Non-Infected', 'Infected'), values = c('#0000FF', '#f09d1d'),
                                             name = leyendText) +
                          theme_bw() + 
                          theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14),
                                legend.text = element_text(size = 14), legend.title = element_text(size = 16),
                                legend.text.align = 0) +
                          scale_x_discrete(labels = c('Female', 'Male')) +
                          theme(legend.position = 'none')
summerBCIplot2 = summerBCIplot + annotate('text',  x = Inf, y = Inf, label = 'A', vjust = 1.5, hjust = 1.5) +
                                   scale_y_continuous(breaks = round(c(seq(-.075, 0.050, 0.025)), 3),
                                                      limits = c(-.060, 0.050))
```

```{r bciplotsave, eco=FALSE}
# Saving BCI plot as pdf
pdf('bcipredictedinfection.pdf')
grid.arrange(summerBCIplot2, winterBCIplot2 , ncol = 2)
dev.off()
grid.arrange(summerBCIplot2, winterBCIplot2, ncol = 2)
```

### Model2. Probability of infection predicted by CV, season, svl and sex

```{r echo=FALSE}
# Training model for infection status prediction
infectionStatusModel = glm(infection.status ~ CV + Season + Sex + SVL, data = matureErythrocyteSeasonalDF, family = binomial(logit))
summary(infectionStatusModel)
```

```{r echo=FALSE}
# Creating empty data frame for infection status prediction
infectionStatusPredictorCVSexSeasonSvl= function(df, model, cv, sex, season, svl){
  # New data frame MUST have model parameters as column names so predict works
  dfBuffer = data.frame(CV = cv, Sex = sex, Season = season, SVL = svl) 

  # Predicting data
  predictedData = predict(model, newdata = dfBuffer, type = 'response', se.fit = T)
  # Generating upper and lower CI
  lower = predictedData$fit - 1.96 * predictedData$se.fit
  upper = predictedData$fit + 1.96 * predictedData$se.fit
  # Data frame for plot
  newDataframe = data.frame(cv, sex, season, svl, predictedData$fit, lower, upper, predictedData$se.fit)
  names(newDataframe) = c('genCV', 'Sex', 'Season', 'SVL', 'predPI', 'lower', 'upper', 'se.predict')
  df = rbind(df, newDataframe)
  return(df)
}
```

```{r echo=FALSE}
# Function to return data from specified sex and season
queryDataFrameData = function(df, sex, season)
{
  return(df[matureErythrocyteSeasonalDF$Sex == sex & matureErythrocyteSeasonalDF$Season == season])
}

# Creating empty data frame for infection status prediction
logitModData = setNames(data.frame(matrix(ncol = 8, nrow = 0)),
                       c('genCV', 'Sex', 'Season', 'SVL', 'predPI', 'lower', 'upper', 'se.predict'))

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'M', 'winter')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = infectionStatusPredictorCVSexSeasonSvl(
  logitModData, infectionStatusModel, CVFWaxis , 'M' , 'winter',
  median(matureErythrocyteSeasonalDF$SVL[matureErythrocyteSeasonalDF$Sex == 'M' & matureErythrocyteSeasonalDF$Season == 'winter'])
)

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'M', 'summer')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = infectionStatusPredictorCVSexSeasonSvl(
  logitModData, infectionStatusModel, CVFWaxis , 'M' , 'summer',
  median(matureErythrocyteSeasonalDF$SVL[matureErythrocyteSeasonalDF$Sex == 'M' & matureErythrocyteSeasonalDF$Season == 'summer'])
)

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'F', 'winter')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = infectionStatusPredictorCVSexSeasonSvl(
  logitModData, infectionStatusModel, CVFWaxis , 'F' , 'winter',
  median(matureErythrocyteSeasonalDF$SVL[matureErythrocyteSeasonalDF$Sex == 'F' & matureErythrocyteSeasonalDF$Season == 'winter'])
)

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'F', 'summer')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = infectionStatusPredictorCVSexSeasonSvl(
  logitModData, infectionStatusModel, CVFWaxis , 'F' , 'summer',
  median(matureErythrocyteSeasonalDF$SVL[matureErythrocyteSeasonalDF$Sex == 'F' & matureErythrocyteSeasonalDF$Season == 'summer'])
)
```

```{r, echo=FALSE}
knitr::kable(logitModData[1:20, ], caption = 'Predicted probability of infection by the coefficient of variation (CV)')
```

### Figure 2

Predicted relationship between the probability of infection and coefficient of variation (CV) and for female and male *A.gundlachi* during the A) 2015 summer season and B) 2016 winter season. Points represent *A.gundlachi* individual data relating infection status to the coefficient of variation (CV) by snout to vent length (SVL). 

```{r echo=FALSE, out.width='60%'}
# Function to filter data by season
filterDFBySeason = function(df, season) {
  filteredDataFrame = df %>% filter(Season == season)
  return (filteredDataFrame)
}
winterData = filterDFBySeason(logitModData, 'winter')
summerData = filterDFBySeason(logitModData, 'summer')

originalWinter = filterDFBySeason(matureErythrocyteSeasonalDF, 'winter')
originalSummer = filterDFBySeason(matureErythrocyteSeasonalDF, 'summer')

infectionStatusWinterplot <- ggplot(winterData, aes(x = genCV, y = predPI, color = Sex)) +
                                geom_line() + 
                                theme(axis.text = element_text(size = 12),
                                      axis.title = element_text(size = 12, face = 'bold')) +
  
                                geom_ribbon(aes(ymin = lower, ymax = upper, fill = Sex),
                                            colour = NA, linetype = 2, alpha = 0.3) + 
                                labs(x = 'Coefficient of Variation', y = 'Probability of Infection') + 
                                new_scale_color() + 
                                geom_point(data = originalWinter, 
                                           aes(size = SVL, x = CV, y = as.numeric(as.character(infection.status)),
                                               color = as.factor(Size)), show.legend = FALSE, shape = 1) +
                                theme(legend.justification = c(0, -0.1)) +
                                labs(labels = c('Female', 'Male')) +
                                theme_bw() +
                                annotate('text',  x = Inf, y = Inf, label = 'A', vjust = 1.5, hjust = 1.5)

infectionStatusSummerplot <- ggplot(summerData, aes(x = genCV, y = predPI, color = Sex)) +
                                    geom_line(show.legend = FALSE) +
                                    theme(axis.text = element_text(size=12),
                                          axis.title=element_text(size=12, face='bold')) +
                                    geom_ribbon(aes(ymin = lower, ymax = upper, fill = Sex),
                                                colour = NA, alpha = 0.3, show.legend = FALSE) +
                                    labs(x = 'Coefficient of Variation', y = 'Probability of Infectionn') +
                                    theme_bw() +
                                    new_scale_color() +
                                    geom_point(data = originalSummer,
                                               aes(size = SVL, x = CV, y = as.numeric(as.character(infection.status)),
                                                   color = as.factor(Size)), shape = 1) +
                                    theme(legend.justification = c(0, 1.15)) +
                                    scale_color_discrete(guide = 'none') +
                                    theme_bw() +
                                    annotate('text',  x = Inf, y = Inf, label = 'B', vjust = 1.5, hjust = 1.5)

pdf('infectionppredictedbyCV.pdf')
grid.arrange(infectionStatusWinterplot, infectionStatusSummerplot, nrow = 2)
dev.off()
grid.arrange(infectionStatusWinterplot, infectionStatusSummerplot, nrow = 2)
```

Females and males of both summer and winter season displayed an increasing relationship between the probability of infection and the coefficient of variation of erythrocytes. SVL had an estimate of 0.072 \textpm 0.02, while CV has an estimate of 0.36 \textpm  0.12.

### Model3. Body condition index (BCI) predicted by CV, season, and sex

```{r echo=FALSE}
# Training model for body condition index prediction
BCIModelwithCV = glm(BCI ~ CV + Season + Sex, data = matureErythrocyteSeasonalDF)
summary(BCIModelwithCV)

BCIPredictorCVSexSeason= function(df, model, cv, sex, season){
  # Makes a new data frame which MUST have model parameters as column names so predict works
  bufferDF = data.frame(CV = cv, Sex = sex, Season = season) 

  # Predicting data
  predictedData = predict(model, newdata = bufferDF, type = 'response', se.fit = T)
  #Generating uper and lower CI
  lower = predictedData$fit - 1.96 * predictedData$se.fit
  upper = predictedData$fit + 1.96 * predictedData$se.fit
  # Data frame for plot
  newDataframe = data.frame(cv, sex, season, predictedData$fit, lower, upper, predictedData$se.fit)
  names(newDataframe) = c('genCV', 'Sex', 'Season', 'predBCI', 'lower', 'upper', 'se.predict')
  
  df = rbind(df, newDataframe)
  return(df)
}
```

```{r echo=FALSE}
# Creating empty data frames for BCI prediction
columnTitlesWithCV = c('genCV', 'Sex', 'Season', 'predBCI', 'lower', 'upper', 'se.predict')
logitModData = setNames(data.frame(matrix(ncol = 8, nrow = 0)), columnTitlesWithCV)

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'M', 'winter')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = BCIPredictorCVSexSeason(logitModData, BCIModelwithCV, CVFWaxis , 'M' , 'winter' )

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'M', 'summer')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = BCIPredictorCVSexSeason(logitModData, BCIModelwithCV, CVFWaxis , 'M' , 'summer' )

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'F', 'winter')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = BCIPredictorCVSexSeason(logitModData, BCIModelwithCV, CVFWaxis, 'F', 'winter')

CVFW  = queryDataFrameData(matureErythrocyteSeasonalDF$CV, 'F', 'summer')
CVFWaxis = seq(floor(min(CVFW)), ceiling(max(CVFW)), length.out = length(CVFW))

logitModData = BCIPredictorCVSexSeason(logitModData, BCIModelwithCV, CVFWaxis, 'F', 'summer')
```

```{r, echo=FALSE}
winterData = filterDFBySeason(logitModData, 'winter')
summerData = filterDFBySeason(logitModData, 'summer')

originalWinter = filterDFBySeason(matureErythrocyteSeasonalDF, 'winter')
originalSummer = filterDFBySeason(matureErythrocyteSeasonalDF, 'summer')
```

### Figure 3

Predicted relationship between the coefficient of variation (CV) and body condition index (BCI) for female and male *A.gundlachi* during the A) 2015 summer season and B) 2016 winter season. Points represent *A.gundlachi* individual data relating the coefficient of variation (CV) to body condition index (BCI).

```{r, echo=FALSE}
# Plotting winter predicted partial relationships between CV and BCI
cvWinterPlot <- ggplot(winterData,
                       aes(x = genCV, y = predBCI, color = Sex)) +
                       geom_line() +
                       theme(axis.text = element_text(size = 12),
                             axis.title = element_text(size = 12, face = 'bold')) +
                       geom_ribbon(aes(ymin = lower, ymax = upper, fill = Sex),
                                   colour = NA, linetype = 2, alpha = 0.3) +
                       labs(x = 'Coefficient of Variation', y = 'Body Condition Index ') +
                       annotate('text',  x = Inf, y = Inf, label = 'A', vjust = 27, hjust = 1) +
                       new_scale_color() +
                       geom_point(data  = originalWinter, aes(x = CV, y = BCI, size=SVL),
                                  show.legend = FALSE, shape = 1) +
                       theme(legend.justification = c(0, -.1)) +
                       labs(labels = c('Female', 'Male')) +
                       theme_bw() +
                       scale_size(range = c(0, 4))

cvSummerPlot <- ggplot(summerData,
                       aes(x = genCV, y = predBCI, color = Sex)) +
                       geom_line(show.legend = FALSE) +
                       theme(axis.text = element_text(size = 12),
                             axis.title = element_text(size = 12, face = 'bold')) +
                       geom_ribbon(aes(ymin = lower, ymax = upper, fill = Sex),
                                   colour = NA, alpha = 0.3, show.legend = FALSE) +
                       labs(x = 'Coefficient of Variation', y = 'Body Condition Index') +
                       annotate('text',  x = Inf, y = Inf, label = 'B', vjust = 27, hjust = 1) +
                       theme_bw() +
                       new_scale_color() +
                       geom_point(data = originalSummer, aes(x = CV, y = BCI, size = SVL), shape = 1) +
                       scale_colour_gradient(guide = 'none') +
                       theme(legend.justification = c(0, 1.15)) +
                       scale_size(range = c(0, 4))

pdf('CVandBCI.pdf')
grid.arrange(cvWinterPlot, cvSummerPlot, nrow = 2)
dev.off()
grid.arrange(cvWinterPlot, cvSummerPlot, nrow = 2)
```
